### 2016-08-02 eda/word-cloud.Rmd
wordclouds one/five-star

---
title: "word-cloud"
author: "Anuj Desai"
date: "August 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(wordcloud)
library(readr)
library(dplyr)
library(tidyr)
```

## Word Clouds
```{r}
reviews_text <- read_csv("../clean_data/reviews_text.csv")
one <- reviews_text %>% 
  filter(stars == 1) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)
five <- reviews_text %>% 
  filter(stars == 5) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)

col <- c("salmon1", "tomato2", "orangered3", "red3", "firebrick4")
wordcloud(one$word, one$n, scale = c(4, 0.25),
          random.order = F, random.color = F,
          use.r.layout = T, rot.per = .35,
          colors = col)
green <- c("palegreen", "springgreen2", "mediumseagreen", "forestgreen", "darkgreen")
wordcloud(five$word, five$n, scale = c(4, 0.25),
          random.order = F, random.color = F,
          use.r.layout = T, rot.per = .35,
          colors = green)
```