### 2016-08-03 eda/business-attributes.Rmd
location data explanation added

---
title: "business-attributes"
output: html_document
---

```{r packages, include = FALSE}
library(readr)
library(stringr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggmap)
library(maps)
library(jsonlite)
```

## Where most reviews are clustered in US

Among 70k+ businesses registered on Yelp, I wanted to figure out more about what makes a business successful. Before investigating deeper, I decided to look at businesses that has 300+ reviews on Yelp by users in order to get more objective ratings of the business. I filtered the data set to get rid of closed businesses, businesses outside of US and limited the business type to restaurants.

```{r}
us_business <- read_csv('../clean_data/us-business.csv', stringsAsFactors=FALSE)
noise_level <- read_csv('../clean_data/noise-attr.csv', stringsAsFactors=FALSE) 
attire_type <- read_csv('../clean_dta/attire-attr.csv', stringsAsFactors=FALSE)
states <- map_data('state')
ggplot(states) + 
  geom_polygon(aes(x = long, 
                   y = lat, 
                   group = group), 
               fill = "grey80", 
               color = "white") + 
  coord_fixed(1.3) + 
  geom_point(data = us_business, 
             aes(long, 
                 lat, 
                 color = factor(stars), 
                 size = count), 
             alpha = 0.7) + 
  guides(colour = guide_legend("Star Rating"), 
         size = guide_legend("Review Counts")) + 
  labs(title = "Restaurant Reviews in United States") +
  theme_nothing(legend = TRUE)
```

## Exploring Noise Level

As expected, restaurants with many reviews and/or high ratings were clustered in specific regions. This table shows where most restaurants are clustered around. I have chosen 4 regions with most businesses: Nevada, Arizona, North Carolina and Pennsylvania.

```{r}
top_data <- us_business %>% 
  group_by(state) %>% 
  tally()
top_data
```

There were many attributes per restaurant, and some were more influential directly/indirectly to a business' success. I have first decided to see how noise level would influcence the star ratings of customers. 
I grouped the data frame by 'state' and 'noise' and created a new column having the average star rating per group. The plot didn't know which noise level to plot first, so I have created the factor levels in noise level.

```{r}
ggplot(data=noise_level, aes(x = noise_factored,
                            y = star, 
                            fill = state)) +
  geom_bar(stat = "identity", 
           position = "dodge") +
  scale_x_discrete(labels = c("Quiet", "Average", "Loud", "Very Loud")) +
  labs(title = "How Noise Level Influences Average Star Ratings") + 
  xlab("Noise Level") + 
  ylab("Average Star Ratings") +
  guides(fill = guide_legend("Region"))
```

## Exploring Attire Types
Now to another business attribute: attire. Similar to noise data, attire_data was retrived by grouping the data frame by 'state' and 'attire'. Again, the star ratings were the means of each group's star ratings.

```{r}
ggplot(data=attire_type, aes(x=attire, y=star, fill=state)) +
  geom_bar(stat="identity", position = "dodge") +
  labs(title = "How Attire Influences Average Star Ratings") + 
  xlab("Attire") + 
  ylab("Average Star Ratings") + 
  guides(fill = guide_legend("Region"))
```