### 2016-08-03 eda/word-cloud.Rmd

---
title: "word-cloud"
author: "Anuj Desai"
date: "August 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(wordcloud)
library(readr)
library(dplyr)
library(tidyr)
source("../functions/wordcloud.R")
```

## Word Clouds

Created some word clouds for one-star and five-star businesses to see if there's a difference between the words reviewers use for each rating using __wordcloud__ library. Expect to see a relationship between negative words and one-star ratings and positive words and five-star ratings.
```{r}
reviews_text <- read_csv("../clean_data/reviews_text.csv")
one <- reviews_text %>% 
  filter(stars == 1) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)

five <- reviews_text %>% 
  filter(stars == 5) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)

red <- c("salmon1", "tomato2", "orangered3", "red3", "firebrick4")
word_cloud(one$word, one$n, red)
green <- c("chartreuse2", "mediumseagreen", "forestgreen", 
           "seagreen", "darkgreen")
word_cloud(five$word, five$n, green)
```

The main words are similar for both clouds like "food" and "service", probably because it was bad food and service for one-star restaurants and great food and service for five-star restaurants. Evidence of this is present in the other words like "love", "friendly", "excellent", "happy" for five-stars and "bad", "rude", "worst", "wait", "horrible" for one-stars.