### 2016-08-05 eda/location-ratings.Rmd

---
title: "Location Ratings"
author: "Irene Lee"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggmap)
library(maps)
library(jsonlite)
```

## User Reviews Are Influenced While Traveling

I wanted to see if YELP users' review ratings change when they are traveling. This required 3 different data sets: business data, reviews data and user data. I read three csv's that had previously been created. 
```{r}
users <- read_csv("../clean_data/yelp_users.csv")
business <- read_csv("../clean_data/yelp_business.csv")
reviews <- read_csv("../clean_data/yelp_reviews.csv")
```

In order to show a good representation of user trends, I have chosen users that have left more than 50 reviews in total. I have joined three tables, selected users(review counts over 50), reviews and business locations.

Since I was using an academic data set provided by Yelp, it only contained some amount of user reviews. I have filtered the users to those with more than 10 reviews in the actual data set provided. I also added a column named 'most_reviewed' which will be true for one state that the user has left most reviews in (the state the user is most familiar with).

The final data set was produced by combining everything above with users' average star rating per state.

```{r}
selected_users <- users %>% 
  filter(review_count > 50) %>% 
  select(name, user_id, review_count)

user_reviews <- reviews %>% 
  select(user_id, review_id, stars, business_id)

business_locations <- business %>% 
  select(business_id, state)

user_reviews <- left_join(selected_users, user_reviews) %>% 
  left_join(business_locations) %>% 
  na.omit

familiar_location <- group_by(user_reviews, user_id, state) %>% 
  tally()

location_reviews <- aggregate(n ~ user_id, familiar_location, max) %>%
  left_join(familiar_location) %>% 
  mutate(most_reviewed = TRUE) %>% 
  filter(n > 10) %>% 
  select(-n)

user_reviews <- left_join(user_reviews, location_reviews) 
user_reviews[is.na(user_reviews)] <- FALSE

# average stars for unfamiliar/familiar area for a user
avg_star_reviews <- user_reviews %>% 
  group_by(user_id, most_reviewed) %>% 
  summarise(avg_star = mean(stars))

```

## Plot

The plot has two histograms, one with user ratings on familiar places and another on unfamiliar places. There is a quite noticeable differences between the two overlayed histograms. The data is extracted from more than 2 million reviews by users, so it for sure represents the .

```{r}
ggplot(avg_star_reviews, aes(x = avg_star, 
              fill = most_reviewed)) + 
  geom_density(alpha = 0.5,
               color = NA) +
  labs(title = "Reviewers' Behavior Changes When In Unfamilar Locations") +
  xlab("Average Star Ratings") + 
  ylab("Density") +
  scale_fill_manual(name ="Regions",
                    values = c("blue", "red"),
                    labels = c("Familiar", "Unfamiliar"))
```